Implementation of Adaptive Waveform Inversion via a Penalty Function

### 2023-12-30 repro/papers/awigrad/paper.tex
near done

The acoustic parameter fields, bulk modulus ($\kappa$) and buoyancy ($\beta$) appearing in the system \ref{eqn:awe} are assumed smooth and uniformly bounded.
Since the speed of propagation is bounded as is the set of source locations $Q_s$ and the interval of propagation $[0,t_D]$, the values of $\kappa(\bx)$ and $\beta(\bx)$ for sufficiently lartge $|\bx|$ have no effect on the solution of the system \ref{eqn:awe}, so $\log \kappa$ and $\log \beta$ can be assumed square-integrable without loss of generality. The set of feasible models $M$ is a  bounded open subset of ${\cal M} = C^{\infty}(\bR^3) \times C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3) \times L^{\infty}(\bR^3) \cap L^2(\bR^3) \times L^2(\bR^3)$.

@@

@@

\section{Derivatives}
The fundamental result of variable projection theory
\cite[]{GolubPereyra:73} applies under the conditions presented here:
the stationary points of $J_{\alpha,\sigma}$ and
$\tilde{J}_{\alpha,\sigma}$ are in bijective correspondence. Moreover,
\[
D_m\tilde{J}_{\alpha,\sigma}[m]\delta m D_mJ[u,m]\delta m|_{u =
  u_{\alpha,\sigma}[m,d]}
\]
\begin{equation}
= \langle (D_mS[m]u\delta m)_{u =
  u_{\alpha,\sigma}[m]} (S[m]u_{\alpha,\sigma}[m]-d \rangle + \alpha^2 \langle D_m(T_{\sigma}[m]u)\delta m)_{u=u_{\alpha,\sigma}[m]},T_{\sigma}[m]u_{\alpha,\sigma}[m]\rangle_U.
\label{eqn:basederiv}
\end{equation}
Note that the regularization term does not appear in the VPM gradient expression, since it does not depend explicitly on $m$.

@@

With the notational conventions listed above, the normal equation \ref{eqn:blocknormal} for $u_{0,\sigma}[m]$ implies
\[
  D_mu_{0,\sigma}[m] \delta m = (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\begin{equation}
  \label{eqn:blockderiv}
  \times ((DS[m]^{T}\delta m) (d - S[m]u_{0,\sigma}[m]) - S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\end{equation}
The single-trace version of the definition \ref{eqn:ppf} is
\[
  T_{\sigma}[m]u)(t) = \frac{t u(t)}{\|u_{0,\sigma}[m]\|},\,u \in L^2[-t_U,t_U],
\]
so
\[
D(T_{\sigma}[m] u)\delta m = -\frac{t u}{\|u_{0,\sigma}[m]\|^3}\langle D_mu_{0,\sigma}[m]\delta u, u_{0,\sigma}[m]\rangle.
\]
\[
  =-\frac{t u}{\|u_{0,\sigma}[m]\|^3}\langle  (S[m]^TS[m] +
  \sigma^2I)^{-1}((DS[m]^{T}\delta m) (d - S[m]u_{0,\sigma}[m])
\]
\[
   - S[m]^T(DS[m]\delta m) u_{0,\sigma}[m]), u_{0,\sigma}[m] \rangle
\]
Define
\begin{equation}
  \label{eqn:defvsigma}
  v_{\sigma} = (S[m]^TS[m]+ \sigma^2I)^{-1} u_{0,\sigma}[m] \in
  L^2[-t_U,t_U].
\end{equation}
Then
\[
  D(T_{\sigma}[m] u)\delta m =
  \frac{t u}{\|u_{0,\sigma}[m]\|^3}[\langle (DS[m]\delta
  m)v_{\sigma}[m], S[m]u_{0,\sigma}[m]-d\rangle
\]
\[+
  \langle S[m]v_{\sigma}[m], (DS[m]\delta m) u_{0,\sigma}[m] \rangle ]
\]
and
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle =
\frac{\langle t u, tu \rangle}{\|u_{0,\sigma}[m]\|^4}
\]
\[
\times [\langle (DS[m]\delta
  m)v_{\sigma}[m], S[m]u_{0,\sigma}[m]-d\rangle
+
  \langle S[m]v_{\sigma}[m], (DS[m]\delta m) u_{0,\sigma}[m] \rangle ]
\]
\[
  = \frac {\|t u\|^2}{\|u_{0,\sigma}[m]\|^4}
\]
\[
  \times [\langle DF[m]\delta m, K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{\sigma}[m]\rangle ]
\]

\[
  = \langle DF[m]\delta m, W(RT_{\sigma}u,2,0)W(Ru_{0,\sigma}[m],-2,0)
\]
\begin{equation}
  \label{eqn:penderiv}
  \times ( K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{\sigma}[m])\rangle 
\end{equation}

Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\large[ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\[
  + \alpha^2 W(RT_{\sigma}[m]u_{\alpha,\sigma}[m],2,0)W(Ru_{0,\sigma}[m],-2,0) (
  K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:deriv}
+ K[u_{0,\sigma}[m]]^T S[m]v_{\sigma}[m]) \large]
\end{equation}
with $v_{\sigma}$ defined in equation \ref{eqn:defvsigma}.

@@

The gradient is unlikely to be needed without the value
of $\tilde{J}_{\alpha,\sigma}$, so 
$T_{\sigma}[m]u_{\alpha,\sigma}$ will have been computed and can be
cached. The same is true of the vector of trace norms
$Ru_{0,\sigma}[m]$. So all ingredient for application of
$W(Ru_{0,\sigma}[m],-2,0)W(RT_{\sigma}[m] u_{\alpha,\sigma}[m],2,0)$

\begin{enumerate}
\item given $m$, $d$, $\alpha$, and $\sigma$, compute
  $u_{0,\sigma}[m]$ by solving the normal equations \ref{eqn:normal}
  for $\alpha=0$, and $T_{\sigma}[m]=W(Ru_{0,\sigma},-1,1)$ (this has
  likely already been done to compute the value of $\tilde{J}_{\alpha,\sigma}$;
\item use $S[m]$, $T_{\sigma}[m]$ to compute $u_{\alpha,\sigma}[m]$ by
  solving the normal equations \ref{eqn:normal}, and record
  $T_{\sigma}[m]u_{\alpha,\sigma}[m]$ as a by-product;
\item compute $v_{0\sigma} \in U$ by solving $(S[m]^TS[m] + \sigma^2 I)v_{0,\sigma} = u_{0,\sigma}[m]$
\item compute $r_0 = S[m]u_{0,\sigma}[m]-d,\,r_{\alpha} = S[m]u_{\alpha,\sigma}[m]-d$ (these may be by-products of steps 1 and 3;
\item compute four vectors in $D$:
  \begin{enumerate}
  \item $G_1 = K[u_{\alpha,\sigma}[m]]^T r_{\alpha}$
  \item $G_2 = K[v_{\sigma}[m]]^Tr_0$
  \item $G_3 = K[u_{0,\sigma}[m]]^T S[m]v_{\sigma}[m]$
  \item $G_0 = G_1 + \alpha^2  W(Ru_{0,\sigma}[m],-2,0)W(RT_{\sigma}[m]u_{\alpha,\sigma}[m],2,0) (G_2 + G_3)$
  \end{enumerate}
\item assemble
  $\nabla \tilde{J}_{\alpha,\sigma}[m] = DF[m]^TG_0$
\end{enumerate}

\subsection{Warner-Guasch AWI}
The (``forward'') AWI objective defined in \cite{Warner:16} is
the limit
\[
  \tilde{J}_{\sigma}[m] = \lim_{\alpha \rightarrow 0}
  \frac{1}{\alpha^2}(\tilde{J}_{\alpha,\sigma}[m]-\tilde{J}_{0,\sigma}[m])
\]
\begin{equation}
  \label{eqn:wg}
  = \frac{1}{2} \|T_{\sigma}[m]u_{0,\sigma}\|^2
\end{equation}

\cite{Warner:16} presume that $S[m]u_{0,\sigma} = d$, which is
manifestly not the case for $\sigma > 0$, however for small $\sigma$
the error is presumably negligible. Neglecting it, the gradient becomes
\[
  \nabla \tilde{J}_{\sigma}[m] = DF[m]^T
    W(Ru_{0,\sigma}[m],-2,0)W(RT_{\sigma}[m]u_{0,\sigma}[m],2,0)
\]
\begin{equation}
  \label{eqn:wggrad}
\times K[u_{0,\sigma}[m]]^T S[m](S[m]^TS[m] + \sigma^2I)^{-1}u_{0,\sigma}[m]).
\end{equation}

>>>>

The acoustic parameter fields, bulk modulus ($\kappa$) and buoyancy ($\beta$) appearing in the system \ref{eqn:awe} are assumed smooth and uniformly bounded.
Since the speed of propagation is bounded as is the set of source locations $Q_s$ and the interval of propagation $[0,t_D]$, the values of $\kappa(\bx)$ and $\beta(\bx)$ for sufficiently large $|\bx|$ have no effect on the solution of the system \ref{eqn:awe}, so $\log \kappa$ and $\log \beta$ can be assumed square-integrable without loss of generality. The set of feasible models $M$ is a  bounded open subset of ${\cal M} = C^{\infty}(\bR^3) \times C^{\infty}(\bR^3) \cap L^{\infty}(\bR^3) \times L^{\infty}(\bR^3) \cap L^2(\bR^3) \times L^2(\bR^3)$.

@@

For convenience, note these obvious properties of $W$:
\begin{itemize}
\item[W1. ] $W(\cdot,\cdot,0)$ is block-scalar, hence commutes with any
  block-diagonal operator;
\item[W2. ] $W(q,p_{s,1},p_{t,1})W(q,p_{s,2},p_{t,2}) = W(q,p_{s,1}+p_{s,2},p_{t,1}+p_{t,2})$;
\item[W3. ] $W(q_1,p_s,p_{t,1})W(q_2,p_s,p_{t,2}) = W(q_1,q_2,
  p_s,p_{t,1}+p_{t,2})$
\end{itemize}

@@

\section{Derivatives}
The fundamental result of variable projection theory
\cite[]{GolubPereyra:73} applies under the conditions presented here:
the stationary points of $J_{\alpha,\sigma}$ and
$\tilde{J}_{\alpha,\sigma}$ are in bijective correspondence. Moreover,
\[
D_m\tilde{J}_{\alpha,\sigma}[m]\delta m = D_mJ[u,m]\delta m|_{u =u_{\alpha,\sigma}[m,d]}
\]
\[
  = \langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]} (S[m]u_{\alpha,\sigma}[m]-d) \rangle
\]
\begin{equation}
+ \alpha^2 \langle D_m(T_{\sigma}[m]u)\delta m|_{u=u_{\alpha,\sigma}[m]},T_{\sigma}[m]u_{\alpha,\sigma}[m]\rangle_U.
\label{eqn:basederiv}
\end{equation}
Note that the regularization term does not appear in the VPM gradient expression, since it does not depend explicitly on $m$.

@@

With the notational conventions listed above, the normal equation \ref{eqn:blocknormal} for $u_{0,\sigma}[m]$ implies
\[
  D_mu_{0,\sigma}[m] \delta m = -(S[m]^TS[m] + \sigma^2I)^{-1}
\]
\begin{equation}
  \label{eqn:blockderiv}
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) + S[m]^T(DS[m]\delta m) u_{0,\sigma}[m])
\end{equation}
The single-trace version of the definition \ref{eqn:ppf} is
\[
  T_{\sigma}[m]u)(t) = \frac{t u(t)}{\|u_{0,\sigma}[m]\|},\,u \in L^2[-t_U,t_U],
\]
so
\[
D(T_{\sigma}[m] u)\delta m = -\frac{t u}{\|u_{0,\sigma}[m]\|^3}\langle D_mu_{0,\sigma}[m]\delta u, u_{0,\sigma}[m]\rangle.
\]
\[
  =-\frac{t u}{\|u_{0,\sigma}[m]\|^3}\langle  (S[m]^TS[m] +
  \sigma^2I)^{-1}((DS[m]^{T}\delta m) (d - S[m]u_{0,\sigma}[m])
\]
\[
   - S[m]^T(DS[m]\delta m) u_{0,\sigma}[m]), u_{0,\sigma}[m] \rangle
\]
Define
\begin{equation}
  \label{eqn:defv0tilde}
  \tilde{v}_{0,\sigma}[m] = (S[m]^TS[m]+ \sigma^2I)^{-1} u_{0,\sigma}[m] \in
  L^2[-t_U,t_U].
\end{equation}
Then
\[
  D(T_{\sigma}[m] u)\delta m =
  \frac{t u}{\|u_{0,\sigma}[m]\|^3}[\langle (DS[m]\delta
  m)\tilde{v}_{0,\sigma}[m], S[m]u_{0,\sigma}[m]-d\rangle
\]
\[+
  \langle S[m]\tilde{v}_{0,\sigma}[m], (DS[m]\delta m) u_{0,\sigma}[m] \rangle ]
\]
and
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle =
\frac{\langle t u, tu \rangle}{\|u_{0,\sigma}[m]\|^4}
\]
\[
\times [\langle (DS[m]\delta
  m)\tilde{v}_{0,\sigma}[m], S[m]u_{0,\sigma}[m]-d\rangle
+
  \langle S[m]\tilde{v}_{0,\sigma}[m], (DS[m]\delta m) u_{0,\sigma}[m] \rangle ]
\]
\[
  = \frac {\|t u\|^2}{\|u_{0,\sigma}[m]\|^4}
\]
\[
  \times [\langle DF[m]\delta m, K[\tilde{v}_{0,\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]\tilde{v}_{0,\sigma}[m]\rangle ]
\]

\[
  = \langle DF[m]\delta m, W(R(W(Ru_{0,\sigma}[m],-2,1)u),2,0)
\]
\begin{equation}
  \label{eqn:penderiv}
  \times ( K[\tilde{v}_{0,\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]\tilde{v}_{0,\sigma}[m])\rangle
\end{equation}

It will turn out to be convenient to take advantage of the {\em
  block-scalar} property of the operator
$W(R(W(Ru_{0,\sigma}[m],-2,1)u),2,0)$. The other operators on the RHS
of equation \ref{eqn:penderiv} are all block-diagonal, and the
right-hand side is linear in $\tilde{v}_{0,\sigma}[m]$. Thus the
$W...$ operator can be combined with the definition of $\tilde{v}...$:
define
\[
  v_{0,\sigma}[m,u]=W(R(W(Ru_{0,\sigma}[m],-2,1)u),2,0) \tilde{v}_{0,\sigma}[m]
\]
to be the solution of the system
\begin{equation}
  \label{eqn:defv0}
  (S[m]^TS[m]+ \sigma^2I) v_{0,\sigma}[m,u] = W(R(W(Ru_{0,\sigma}[m],-2,1)u),2,0) u_{0,\sigma}[m].
\end{equation}

Then equation \ref{eqn:penderiv} can be re-written as
\[
  \langle D(T_{\sigma}[m]u)\delta m, T_{\sigma}[m]u\rangle = \langle DF[m]\delta m, 
\]
\begin{equation}
  \label{eqn:penderiv}
   K[v_{0,\sigma}[m,u]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u]\rangle
\end{equation}
Combine equations \ref{eqn:resderiv} and \ref{eqn:penderiv} with
equation \ref{eqn:basederiv} to obtain
\[
  \nabla \tilde{J}_{\alpha,\sigma}[m] =
  DF[m]^T\large\{ K[u_{\alpha,\sigma}[m]]^T (S[m]u_{\alpha,\sigma}[m]-d)
\]
\begin{equation}
  \label{eqn:deriv}
  + \alpha^2 (
  K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+ K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{\alpha,\sigma}[m]]) \large\}.
\end{equation}

@@

The gradient is unlikely to be needed without the value
of $\tilde{J}_{\alpha,\sigma}$, so the vector of trace norms
$Ru_{0,\sigma}[m]$ will have been computed, as it is a key ingredient
in the definition of the penalty
$T_{\sigma}[m]u_{\alpha,\sigma}$, and can be
cached.

\begin{enumerate}
\item given $m$, $d$, $\alpha$, and $\sigma$, compute
  $u_{0,\sigma}[m]$ by solving the normal equations \ref{eqn:normal}
  for $\alpha=0$, also $Ru_{0,\sigma}$ (this has
  likely already been done to compute the value of $\tilde{J}_{\alpha,\sigma}$);
\item compute $u_{\alpha,\sigma}[m]$ by
  solving the normal equations \ref{eqn:normal}, and record
  $R(W(Ru_{0,\sigma}[m],-2,1)u_{\alpha,\sigma}[m])$ as a by-product;
\item compute $v_{0\sigma}[m, u_{\alpha,\sigma}[m]] \in U$ by solving the linear system
  \ref{eqn:defv0};
\item compute $r_0 = S[m]u_{0,\sigma}[m]-d,\,r_{\alpha} = S[m]u_{\alpha,\sigma}[m]-d$ (these may be by-products of steps 1 and 3;
\item compute four vectors in $D$:
  \begin{enumerate}
  \item $G_1 = K[u_{\alpha,\sigma}[m]]^T r_{\alpha}$
  \item $G_2 = K[v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]]^Tr_0$
  \item $G_3 = K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m, u_{\alpha,\sigma}[m]]$
  \item $G_0 = G_1 + \alpha^2 (G_2 + G_3)$
  \end{enumerate}
\item assemble
  $\nabla \tilde{J}_{\alpha,\sigma}[m] = DF[m]^TG_0$
\end{enumerate}

\subsection{Warner-Guasch AWI}
The (``forward'') AWI objective defined in \cite{Warner:16} is
the limit
\begin{equation}
  \label{eqn:wglim}
  \tilde{J}_{\sigma}[m] = \lim_{\alpha \rightarrow 0}
  \frac{1}{\alpha^2}(\tilde{J}_{\alpha,\sigma}[m]-\tilde{J}_{0,\sigma}[m])
\end{equation}
\begin{equation}
  \label{eqn:wg}
  = \frac{1}{2} \|T_{\sigma}[m]u_{0,\sigma}\|^2
\end{equation}

In the following paragraphs, I will compute the $m$ derivative of
$\tilde{J}_{\sigma}$ directly from the right-hand side of the
definition \ref{eqn:wg}. Then I will compute it as the limit of
derivatives as suggested in the identity \ref{eqn:wglim}, and show
that these are the same.  A by-product of this comparison is the
observation that the limit \ref{eqn:wglim} is in the sense of
$C^1(L^{\infty}(\bR^3) \times L^{\infty}(\bR^3))$.

The derivative of the expression on the right-hand side of
\ref{eqn:wg} is the sum of two terms, the first of which is given by
\ref{eqn:penderiv} with $u = u_{0,\sigma}[m]$. The second is
\[
  \langle T_{\sigma}[m]u_{0,\sigma}[m], T_{\sigma}[m]D_m
  u_{0,\sigma}[m]\delta m \rangle
\]
which from identity \ref{eqn:blockderiv} is
\[
  = -\langle  T_{\sigma}[m]^TT_{\sigma}[m]u_{0,\sigma}[m], (S[m]^TS[m] + \sigma^2I)^{-1}
\]
\[
  \times ((DS[m]^{T}\delta m) (S[m]u_{0,\sigma}[m]-d) +
  S[m]^T(DS[m]\delta) m u_{0,\sigma}[m])
\]
Since
$T_{\sigma}[m]^TT_{\sigma}[m] = W(Ru_{0,\sigma}[m],-2,2)$, this is
\[
  = -\langle (DS[m]\delta m) (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m],
  S[m]u_{0,\sigma}[m] - d \rangle
\]
\begin{equation}
  \label{eqn:term2}
  - \langle S[m] (S[m]^TS[m] +
  \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m], (DS[m]\delta
  m)u_{0,\sigma}[m] \rangle
\end{equation}
Introduce the abbreviation
\begin{equation}
  \label{eqn:defv1}
  v_{1,\sigma}[m] = (S[m]^TS[m] + \sigma^2I)^{-1}W(Ru_{0,\sigma}[m],-2,2)u_{0,\sigma}[m].
\end{equation}
Then the RHS of equation \ref{eqn:term2bis} may be re-written as
\begin{equation}
  \label{eqn:term2bis}
  = -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle.
\end{equation}
Recall that we have just computed the second of two terms, the sum of
which is the derivative of $\tilde{J}_{\sigma}[m]$. The first is
\ref{eqn:penderiv} with $u = u_{0,\sigma}[m]$. Combine the two
terms to obtain
\[
  D\tilde{J}_{\sigma}[m] \delta m = \langle DF[m]\delta m,
\]
\[
  K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]^T(S[m]u_{0,\sigma}[m]-d)
+
K[u_{0,\sigma}[m]]^T S[m]v_{0,\sigma}[m,u_{0,\sigma}[m]]\rangle
\]
\[
  -\langle DF[m]\delta m, K[v_{1,\sigma}[m]]^T (S[m]u_{0,\sigma}[m]-d) + 
  K[u_{0,\sigma}[m]]^T S[m] v_{1,\sigma}[m] \rangle
\]
\[
  = \langle DF[m]\delta
  m,(K[v_{0,\sigma}[m,u_{0,\sigma}[m]]]-K[v_{1,\sigma}[m]])^T(S[m]u_{0,\sigma}[m]-d)
\]
\[
  + K[u_{0,\sigma}[m]]^T S[m] (v_{0,\sigma}[m,u_{0,\sigma}[m]] -
  v_{1,\sigma}[m]) \rangle.
\]
To simplify this expression further, and reduce the computational
complexity of the derivative expression, note that the
operator-valued function $K$ is linear in its argument, according to
its definition \ref{eqn:convdef}. So the above is
\begin{equation}
  \label{eqn:wgderiv}
  = \langle DF[m]\delta m, (K[v_{\sigma}[m]]^T(S[m]u_{0,\sigma}[m]-d) 
  + K[u_{0,\sigma}[m]]^T S[m] v_{\sigma}[m] )\rangle.
\end{equation}
in which
\[
  v_{\sigma}[m] = v_{0,\sigma}[m,u_{0,\sigma}[m]] -  v_{1,\sigma}[m].
\]
The definitions \ref{eqn:defv0} and \ref{eqn:defv1} imply that
$v_{\sigma}[m]$ is the solution of the linear system
\begin{equation}
  \label{eqn:defvsigma}
  (S[m]^TS[m]+ \sigma^2I) v_{\sigma}[m] =
  (W(R(W(Ru_{0,\sigma}[m],-2,1)u_{0,\sigma}),2,0) - W(Ru_{0,\sigma}[m],-2,2))
  u_{0,\sigma}[m].
\end{equation}
That is, computation of $v_{\sigma}[m]$, hence of $D
\tilde{J}_{\sigma}[m]\delta m$, requires the solution of only one additional
linear system, beyond that necessary to compute the value of $\tilde{J}_{\sigma}[m]$.

\noindent {\bf Remark:} The expression \ref{eqn:wgderiv} immediately
yields a formal prescription for the gradient of
$\tilde{J}_{\sigma}[m]$ as the output of the adjoint modeling operator
$DF[m]^T$ applied to an ``adjoint source'' (the right-hand side of the
inner product \ref{eqn:wgderiv}. \cite{Warner:16} point out this
similarity with FWI, in which the adjoint source (in the simplest
version) is the residual $F[m]-d$. Consequently, FWI implementations
already contain a great deal of the code necessary to implement
AWI. \cite{Warner:16} also presume that the regularization weight
$\sigma$ is sufficiently small that the adaptive filter $u_{0,\sigma}$
effectively fits the data. Thus $S[m]u_{0,\sigma}-d$ is negligible,
regardless of $m$, and the first term on the right-hand side of the
inner product \ref{eqn:wgderiv} drops out, further simplifying the AWI
adjoint source.

It remains to establish that the same result (expression
\ref{eqn:wgderiv}) for the derivative can be derived from the
limit \ref{eqn:wglim}. From from equation \ref{eqn:basederiv},
\[
  \frac{1}{\alpha^2}(D \tilde{J}_{\alpha,\sigma}[m] \delta m - D
  \tilde{J}_{0,\sigma}[m]\delta m) =
\]
\[
  = \frac{1}{\alpha^2} \left(\langle D_m(S[m]u)\delta m_{u =  u_{\alpha,\sigma}[m]}, (S[m]u_{\alpha,\sigma}[m]-d) \rangle\right.
\]
\[
  - \left.\langle D_m(S[m]u)\delta m_{u =  u_{0,\sigma}[m]}, (S[m]u_{0,\sigma}[m]-d) \rangle\right)
\]
\[
  +  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{\alpha,\sigma}[m]},T_{\sigma}[m]u_{\alpha,\sigma}[m]\rangle_U
\]
Since the normal operator \ref{eqn:normal} is uniformly positive
definite hence boundedly invertible uniformly in $\alpha \ge 0$,
$u_{\alpha,\sigma}[m] \rightarrow u_{0,\sigma}[m]$ in $U$. All of the
operators appearing in the third term are $U-$continuous, so  the
third term tends to
\[
  \langle D_m(T_{\sigma}[m]u)\delta
  m|_{u=u_{0,\sigma}[m]},T_{\sigma}[m]u_{0,\sigma}[m]\rangle_U
\]
which is identixal to the left-hand side of equation
\ref{eqn:penderiv} and is the first term in the direct computation of
the $\tilde{J}_{\sigma}$ derivative, analyzed above.
